---
title: "MATH208-A4"
author: "Christopher Zheng. ID: 260760794"
date: "15/11/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
ceramic_data<-dget("ceramic.txt")
```


(a) What class of object is ceramic_data?
```{r}
class(ceramic_data)
```

(b) Using the ceramic_data object, compute the median Mean_Strength by Batch, i.e. write a line of
code that produces a vector that computes the mean of all observations in each batch, respectively.
What is the difference in mean by Batch?
```{r}
# Batch 1
# avg_batch_1 <- sum(ceramic_data[1:2,1:2,1:2,1:2,1:1][1:16]) / 16
# Batch 2
# avg_batch_2 <- sum(ceramic_data[1:2,1:2,1:2,1:2,2:2][1:16]) / 16
# difference <- abs(avg_batch_1 - avg_batch_2)
# avg_vector <- c(avg_batch_1,avg_batch_2)
# print(avg_vector)  # the vector of the two averages
# print(difference)  # the absolute difference between two mean values
m_by_Batch <- apply(ceramic_data, c(5), median)
print(m_by_Batch)
diff_by_Batch <- unname(m_by_Batch[1] - m_by_Batch[2])
print(diff_by_Batch)
```

(c) Compute the overall standard deviation of the Mean_Strength values across all 32 observations.
Hint: remember that this type of object is stored internally as an generic vector.
```{r}
all_obs <- c(ceramic_data[1:2,1:2,1:2,1:2,1:2][1:32])
std_dev <- sd(all_obs)
print(std_dev)
```

(d) Using the ceramic_data object, compute the average Mean_Strength for each of the four groups
defined by the cross-classification of Table_Speed and Direction. Hint: the most succinct way to do
this would return a 2x2 table with Table_Speed levels in the rows and Direction levels in the columns.

```{r}
#sum_table <- ceramic_data[,1,1,,1] + ceramic_data[,2,1,,1] + ceramic_data[,1,2,,1] + ceramic_data[,1,1,,2] + 
#  ceramic_data[,1,2,,2] + ceramic_data[,2,2,,1] + ceramic_data[,2,1,,2] + ceramic_data[,2,2,,2]
#output_table <- sum_table / 8
output_table <- apply(ceramic_data, c(1,4), mean)
print(output_table)
```

(e) Compute the difference in means between the Longitudinal and Transverse levels of Direction for each
level of Slow and Fast separately. Hint: Use part (d) and the apply function.
```{r}
# helper function to calculate the difference of the means and to return the absoluate value of the difference
abs_diff <-function(x){
  return(abs(x[[1]]-x[[2]]))
}
# apply the difference function horizontally
apply(output_table,1,abs_diff)
```



















