---
title: "Math208-A2"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

MATH 208 - Assignment 2 - Christopher Zheng - 206760794


Question 1:

```{r}
#install.packages("fivethirtyeight")
library(fivethirtyeight)
library(tidyverse)
library(ggplot2)

df <- tbl_df(biopics)

```

(a) Using the plot of your choice, assess whether the total number of biopics released per year has increased over time based on the data collected from the IMDB movie database.

```{r}

df_per_year <- df %>% group_by(year_release) %>% summarise(count=n())

ggplot(df_per_year,aes(x=year_release,y=count,group=1)) +
  geom_point() + geom_line(stat="summary",fun.y=mean)

#In conlcusion, the total number of biopics released per year has increased over time based on the timeline.
```


(b) Produce a stacked barplot similar to the barplot in the original article showing the relative numbers of male and female subjects over time (Note the figures will not exactly be the same as the data in the article figures is not the same as in the dataset).

```{r}
df_per_year <- df %>% group_by(year_release) %>% mutate(count=n())

ggplot(df_per_year,aes(x=year_release,fill=subject_sex)) +
  geom_bar() + scale_fill_viridis_d() 


head(df_per_year)
```




(c) Produce a stacked barplot similar to the barplot in the original article showing the relative numbers of white subjects, subjects who are persons of color, and unknown race subjects over time. (Mote the figures will not exactly be the same as the data in the article figures is not the same as in the dataset).
```{r}
df_per_year <- df %>% group_by(year_release) %>% mutate(count=n()) %>%
  mutate(subject_race = ifelse(is.na(subject_race),"Unknown race (NA)",ifelse(subject_race=="White",subject_race,"People of colour")) )
#df_per_year
ggplot(df_per_year,aes(x=year_release,fill=subject_race)) +
  geom_bar() + scale_fill_viridis_d() 

```




(d) Based on a mosaic plot (collapsing over year of release), which sex / white-nonwhite-NA group is the most underrepresented in biopics based on number of subjets?
```{r}
library(ggmosaic)
df_per_year <- df %>% group_by(year_release) %>% mutate(count=n()) %>%
  mutate(subject_race = ifelse(is.na(subject_race),"Unknown race (NA)",ifelse(subject_race=="White",subject_race,"People of colour")) )


ggplot(df_per_year) + 
  geom_mosaic(aes(x=product(year_release), fill=interaction(subject_sex,subject_race)))+
  xlab("Year") + ylab("")

# Based on the mosaic plot, females who are people of colours are the most underrepresented.

```

(e) Produce a summary table containing counts and proportions of biopic subjects per year for each sex/white-nonwehite-NA factor combination.

```{r}
library(dplyr)

df_per_year <- df %>% group_by(year_release) %>% mutate(count=n()) %>%
  mutate(subject_race = ifelse(is.na(subject_race),"Unknown race (NA)",ifelse(subject_race=="White",subject_race,"People of colour")) ) %>%
  select(year_release,subject_race,subject_sex,count)
#df_per_year
df_summary <- df_per_year %>% group_by(year_release,subject_sex,subject_race) %>% mutate(number = n(),proportion=number/count)



df_summary <- df_summary[order(df_summary$year_release),] %>% unique(.) %>% select(year_release,subject_sex,subject_race,number,proportion)
df_summary

```



(f) Create (i) a line plot showing the counts of these groups over time and (ii) a line plot showing the relative proportions of subjects over time. Would you infer from these plots that the imbalance is improving over time or not? Explain your answer.

```{r}
#df_summary

ggplot(df_summary,aes(x=year_release,y=number,fill=interaction(subject_sex,subject_race))) +
  geom_point(aes(color=interaction(subject_sex,subject_race))) + geom_line(stat="summary",fun.y=mean,aes(color=interaction(subject_sex,subject_race))) + 
  xlab("6 in 1 line plot of counts")

ggplot(df_summary,aes(x=year_release,y=number,fill=interaction(subject_sex,subject_race))) +
  geom_point(aes(color=interaction(subject_sex,subject_race))) + geom_line(stat="summary",fun.y=mean,aes(color=interaction(subject_sex,subject_race))) + facet_wrap(~interaction(subject_sex,subject_race))

ggplot(df_summary,aes(x=year_release,y=proportion,fill=interaction(subject_sex,subject_race))) +
  geom_point(aes(color=interaction(subject_sex,subject_race))) + geom_line(stat="summary",fun.y=mean,aes(color=interaction(subject_sex,subject_race))) + 
  xlab("6 in 1 line plot of proportions")

ggplot(df_summary,aes(x=year_release,y=proportion,fill=interaction(subject_sex,subject_race))) +
  geom_point(aes(color=interaction(subject_sex,subject_race))) + geom_line(stat="summary",fun.y=mean,aes(color=interaction(subject_sex,subject_race))) + facet_wrap(~interaction(subject_sex,subject_race))


# The problem of imbalance is getting better as we can tell from the line plot of the proportions that the percentages are all gradually converging to 1/6.

```


Question 2
(a) First, create a summary table that finds the mean and median for each of the six quantitive variables with a column for each group. (Hint: use summarise, pivot_longer, and pivot_wider). Which varible(s) seem to differentiate amongst the different types of diabetes?
```{r}
library(heplots)
df <- tbl_df(Diabetes)
df_grouped <- df %>% group_by(group) %>% summarise_all(list(Avg=mean,Med=median)) %>%
  pivot_longer(cols=contains('_'),names_to = "Measure") %>%
  pivot_wider(id_cols = Measure, names_from = group) %>% arrange(desc(Measure))

# variables that seem to differentiate amongst the different types of diabetes: sspg, instest, glutest, glufast.	

```


(b) Create 3 scatterplots, comparing all possible pairs of the glucose test variable, the insulin test varible and the sspg variable. Which pair of variables seems to allow for the strongest distinction amongst the three groups?
```{r}
#glucose, insulin, sspg


library(heplots)
df <- tbl_df(Diabetes)
df_grouped <- df %>% group_by(group)
ggplot(df_grouped,aes(x=glutest,y=instest,col=group)) + geom_point() +
  labs(x="glutest",y="instest",title="Glucose test vs. Insulin test")

ggplot(df_grouped,aes(x=glutest,y=sspg,col=group)) + geom_point() +
  labs(x="glutest",y="sspg",title="Glucose test vs. sspg")

ggplot(df_grouped,aes(x=instest,y=sspg,col=group)) + geom_point() +
  labs(x="instest",y="sspg",title="Insulin test vs. sspg")

# Glucose test & sspg seem to allow for the max distinction because the points are more clustered in this case.
```



(c) Using the pair of variables that you chose in part (b), make 2-d histograms and contour plots for each group separately. Do you find for this dataset that these plot provide useful summaries of the differences in distributions in the three groups? Feel free to adjust the amount of binning/smoothing and the number of levels from the defalut levels.

```{r}
df_grouped_normal <- df_grouped[df_grouped$group == "Normal",]
df_grouped_chemical <- df_grouped[df_grouped$group == "Chemical_Diabetic",]
df_grouped_overt <- df_grouped[df_grouped$group == "Overt_Diabetic",]

ggplot(df_grouped_normal,aes(x=glutest,y=sspg)) + geom_bin2d(bins=15)+
  scale_fill_continuous(type="viridis") +
  labs(x="glucose test",y="sspg")

ggplot(df_grouped_chemical,aes(x=glutest,y=sspg)) + geom_bin2d(bins=15)+
  scale_fill_continuous(type="viridis") +
  labs(x="glucose test",y="sspg")

ggplot(df_grouped_overt,aes(x=glutest,y=sspg)) + geom_bin2d(bins=15)+
  scale_fill_continuous(type="viridis") +
  labs(x="glucose test",y="sspg")

ggplot(df_grouped_normal,aes(x=glutest,y=sspg)) + geom_density_2d(col="red") +
  labs(x="glucose test",y="sspg") + ylim(c(-100,350)) + xlim(c(0,500))

ggplot(df_grouped_chemical,aes(x=glutest,y=sspg)) + geom_density_2d(col="red") +
  labs(x="glucose test",y="sspg") + ylim(c(0,350)) + xlim(c(300,650))

ggplot(df_grouped_overt,aes(x=glutest,y=sspg)) + geom_density_2d(col="red") +
  labs(x="glucose test",y="sspg") #+ ylim(c(400,1000)) + xlim(c(0,500))


print(df_grouped_overt)
```





